---
title: "Simple Linear Regression"
---


## Introduction

- Statistical model is quite important across a wide range of fields, providing researchers with tools for both explanation and prediction. 
- The most popular models of the statistical practice has been the general linear model (GLM). 
- The GLM finds the relation of a dependent and several independents variables that take form of tools as analysis of variance (ANOVA) and regression.


## Simple Linear Regression

- The simple linear regression model in population form is where $y_i$ is the dependent variable for individual $i$ in the data set and $x_i$ is the independent variable for subject $i$ ($i$ = 1, ..., $N$). 
- The terms $\beta_0$ dan $\beta_1$, are the intercept and slope of the model, respectively. 

$$
y_{ij}=\beta_{0j}+\beta_{1j}x+\varepsilon_{ij}
$$ {#eq-slr}

- The intercept is the point at which the line in @eq-slr crosses the $y$ axis at $x$ = 0. 
- Thus, larger values of $\beta_1$ (positive or negative) indicate a stronger linear relationship between $y$ and $x$.


## Random Error

- Random error, represented by $\varepsilon_i$, is inherent in any statistical model, including regression. 
- It expresses the fact that for any individual, $i$, the model will not generally provide a perfect predicted value of $y$, denoted $\hat{y}_i$ and obtained by applying the regression model as

$$
\hat{y}_i = \beta_0 + \beta_1x_i
$$ {#eq-re}

- Conceptually, this random error is representative of all factors that may influence the dependent variable other than $x$. 


## Estimating Regression with OLS

- Ordinary least squares (OLS) is popular methods for obtaining estimated values of the regression model parameters ($b_0$ and $b_1$, respectively) given a set of $x$ and $y$. 
- $\beta_0$ and $\beta_1$ must be estimated using sample data taken from the population. 
- The goal of OLS is to minimize the sum of the squared differences between the observed values of $y$ and the model-predicted values of $y$, across the sample. 
- This difference, known as the residual, is written as

$$
e_i = y_i - \hat{y}_i
$$ {#eq-diff}


- Therefore, the method OLS seeks to minimize

$$
\Sigma_{i=1}^n e_i^2 = \Sigma_{i=1}^n (y_i - \hat{y}_i)
$$ {#eq-ols}


## OLS Criteria

- It should be noted that in the context of simple linear regression, the OLS criteria reduce to the following equations, which can be used to obtain $b_0$ and $b_1$ as

$$
b_1 = r \left(\frac{S_y}{S_x} \right)
$$ {#eq-olsc1}

and

$$
b_0 = \overline{y} - b_1\overline{x}
$$ {#eq-olsc2}



## Example of Simple Regression

- In this example, we use the PISA 2022 data to analyze the impact of ESCS on the math achievement of Indonesian students. 
- The sample includes 1329 students who were assessed for both variables. 
- In this scenario, math achievement serves as the dependent variable, while ESCS is the independent variable. 
- Descriptive statistics for each variable, along with the correlations between them, are provided in Table 1.1.

## Descriptive stat

```{r}
library(tidyverse)
library(gt)
library(psych)
```


```{r}
library(tidyverse)
pisa_idn <- read_csv("data/pisa_idn_sample.csv") %>% 
  select(ESCS, MATH)

describe(pisa_idn)
```


## Correlation

```{r}
#| echo: false
#| output: false
cor(pisa_idn)
```


```{r}
Variable <- c("Math", "ESCS")
Mean <- c(367.28, -1.63)
Sd <- c(58.34, 0.99)
Correlation <- c(0.244, "")

table <- data.frame(Variable, Mean, Sd, Correlation)

table %>% 
  gt() %>%
  cols_label(
    Variable = md("**Variable**"),
    Mean = md("**Mean**"),
    Sd = md("**Standar Deviasi**"),
    Correlation = md("**Correlation**"))
```

## Beta 1

- Using equations (1.4) and (1.5), we can use this information to obtain estimates for both the slope and the intercept of the regression model. 
- First, the slope of the regression is calculated as

$$
b_1 = 0.244 \left(\frac{58.34}{0.99} \right)=14.38
$$ {#eq-olsc1}

## Beta 0

- The results indicate that individuals with higher ESCS scores generally achieve higher math scores. - We can calculate an estimate of the intercept using the values in the table:

$$
b_0=367.28-(14.38)(-1.63)=390.72
$$

## Full model

- The resulting estimated regression equation for math and ESCS is:

$$
\hat {math}=390.72+14.38(ESCS).
$$

- This indicates that for a 1-point increase in ESCS score, math achievement would increase by 390.72 points.


## Measure the strength

- To assess the strength of the relationship between ESCS and math achievement, we should calculate the coefficient of determination. 
- This requires the values of $SS_R$ and $SS_T$. 
- The sum of squares due to regression ($SS_R$) or explained sum of squares (ESS) is the sum of the differences between the predicted value and the mean of the dependent variable. 

## Sum of squares due to regression 

- We can calculate the strength by this equation.

$$SS_R=\Sigma^n_{i=1}(\hat{y}_i-\bar{y})^2$$ 

- Where $\hat{y}_i$ is the predicted value of the dependent variable and $\bar{y}$ is mean of the dependent variable.



## Sum of squares error

- The sum of squares error ($SS_E$) or residual sum of squares is the difference between the observed and predicted values.

$$SS_E=\Sigma^n_{i=1}\varepsilon^2_i$$ 
Where $\varepsilon_i$ is the difference between the actual value of the dependent variable and the predicted value:

$$\varepsilon_i=y_i-\hat{y}_i$$

## Sum of squares total

- The sum of squares total ($SS_T$) or the total sum of squares (TSS) is the sum $SS_R$ and $SS_E$. 

$$SS_T=\Sigma^n_{i=1}(\hat{y}_i-\bar{y})^2 +\Sigma^n_{i=1}\varepsilon^2_i$$ 

## Regression in R

To simplify the calculation we will calculate the $R^2$ value using r. First, we create a regression model using the lm () function.

```{r}
#| echo: true
model <- lm(MATH ~ ESCS, data = pisa_idn)
```

## Calculate by hand

- We calculate the $SS_R$, $SS_E$, $SS_T$, $R^2$ value with the following command:

```{r}
#| echo: true

ssr <- sum((fitted(model)-mean(pisa_idn$MATH))^2)
ssr

sse <- sum((fitted(model) - pisa_idn$MATH)^2)

sst <- ssr+sse
sst
```

## Calculate R square

$$
R^2=\frac{SS_R}{SS_T}= \frac{269424.3}{4520484}=0.06
$$

- The results indicate that about 6% of the difference in math achievement can be accounted for by the variance in ESCS scores. 

## Calculate F

- With this $R^2$ value, we can compute the F-statistic to test if any of the model slopes (in this instance, there is only one) are different from 0 in the population.

$$
F= (\frac {1329-1-1} {1})(\frac{0.06}{1-0.06})= 84.7
$$

















