---
title: "Hands-on Random Slope"
---


```{r library}
#| echo: false

library(lme4)
library(sjPlot)
library(lattice)
library(haven)
library(tidyverse)
pisa_idn <- read_csv("data/pisa_idn.csv")

skor <- pisa_idn %>%
  filter(CNTSCHID == 36000271 | 
        CNTSCHID == 36000272 |
        CNTSCHID == 36000273 |
        CNTSCHID == 36000274 |
        CNTSCHID == 36000275 |
        CNTSCHID == 36000276 |
        CNTSCHID == 36000277 |
        CNTSCHID == 36000278 |
        CNTSCHID == 36000279 |
        CNTSCHID == 36000280 |
        CNTSCHID == 36000285 |
        CNTSCHID == 36000286 |
        CNTSCHID == 36000287 |
        CNTSCHID == 36000288 |
        CNTSCHID == 36000289 |
        CNTSCHID == 36000290 |
        CNTSCHID == 36000291 |
        CNTSCHID == 36000292 |
        CNTSCHID == 36000293 |
        CNTSCHID == 36000294) 

skor <- skor %>% filter(growth != "NA")
```

# Random Slope

- The previous model makes the assumption that the effect of growth mindset is the same across all schools (that's why the lines are parallel).

- What if one growth mindset score is more ‘effective’ in some schools than others in improving maths scores.

- At this point a random slope is needed.

- We get a new coefficient that describes the differences between schools in the effect of growth mindset on maths scores.

## Random Slope Estimation

```{r}
# model with random slope
m2 <- lmer(MATH ~ 1 + growth + 
             (1 + growth | CNTSCHID), 
           data = skor)
# print results
summary(m2)
```

## Random Slope Estimation

```{r}
tab_model(m2, show.icc = TRUE)
```

## Interpretation

- The fixed effect of growth mindset is smaller than in model 1 and we have a new random effect coefficient.
- The random slope variance for years of education is 73.29.

## Density

```{r}
#| output: false

skor$m2 <- predict(m2)
# visualize the predictions based on our model
skor %>% 
  ggplot(aes(growth, m2)) + 
  geom_smooth(se = F, method = lm, size = 2) +
  geom_jitter()+
  stat_smooth(aes(color = CNTSCHID, group = CNTSCHID),
              geom = "line", alpha = 0.4, size = 1) +
  theme_bw() +
  guides(color = F) +
  labs(x = "Growth Mindset", 
       y = "Math score", 
       color = "School")

```

## Density Result

```{r results='hold'}
#| echo: false

skor$m2 <- predict(m2)
# visualize the predictions based on our model
skor %>% 
  ggplot(aes(growth, m2)) + 
  geom_smooth(se = F, method = lm, size = 2) +
  geom_jitter()+
  stat_smooth(aes(color = CNTSCHID, group = CNTSCHID),
              geom = "line", alpha = 0.4, size = 1) +
  theme_bw() +
  guides(color = F) +
  labs(x = "Growth Mindset", 
       y = "Math score", 
       color = "School")
```

Fixed effect is -8.054, but each school has a diverse slope.

## Density of Intercept 

```{r}
# another way to see random effect
qqmath(ranef(m2, condVar = TRUE))[[1]]
```

The graph shows how the influence of growth mindset varies across schools.

## Random Effects

```{r}
coefs_m2 <- coef(m2)

coefs_m2$CNTSCHID %>%
  mutate(CNTSCHID = rownames(coefs_m2$CNTSCHID))  %>% 
  ggplot(aes(growth, `(Intercept)`, label = CNTSCHID)) + 
  geom_point() + 
  geom_smooth(se = F, method = lm) +
  geom_label(nudge_y = 0.15, alpha = 0.5) +
  theme_bw() +
  labs(x = "Slope", y = "Intercept")
```

## Interpretation

- The graph shows that some schools have high maths scores but low effects on outcomes and vice versa.

- The blue line represents the relationship between the random intercept and the slope with a coefficient of 0.6. This shows that the smaller the fixed mindset, the higher the maths score.
